---
title: "dwc_mapping"
author: "Dimitri Brosens"
date: "17-3-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
#library(sf)             # coordinate transformation
library(lubridate)      # date transformation
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was corrected in Excel
Muskrat occurrences opened in openRefine
Obsolete columns removed
some columns renamed to DwC term
File exported to csv

```{r}
#input_interim <- read_csv2(file = here:here("datasets", "mica-uvw-occurrences", "data", "raw", "2020_11_18_Vangstgegevens #muskus- en beverratten_NL_2019.csv"))  

input_data <- read.csv(file = here::here("datasets", "favv-ixodida", "data", "interim", "Integrated_output_ticks_ITG.csv"), sep = ',')
```

Preview data:

```{r}
input_data %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% remove_empty("rows")

```

```{r}
input_data %<>%
   clean_names()
```


delete obsolete columns

```{r}
input_data %<>% select(-c("male","female","larva","nymph","id_1"))
```


```{r}
input_data %>% head(n = 5)
```

```{r}
input_data %<>%

         mutate(
           Theileriaequi = na_if(theileriaequi, "negative"),
           `babesiacanis_caballi` = na_if(`babesiacanis_caballi`, "negative"),
           TBE = na_if(tbe, "negative"),
           borrelia = na_if(borrelia, "negative"),
           host = "host"
         ) %>%
         gather(
           parasite,
           value,
           host, theileriaequi, `babesiacanis_caballi`, tbe, borrelia,
           na.rm = TRUE,
           convert = TRUE
         )




```


```{r}

input_data %<>%
   mutate(scientificName = if_else(
   parasite == "host",
   paste(genus, species),
   parasite
             ))


```




# Georeference source data
data is already in WGS84

# Create eventID


```{r}
input_data %<>% mutate(dwc_eventID = str_c("IGT:FAVV:T:EV:",str_replace(collectiondate,",","-") ,str_replace(extractiondate, ",","-") ,str_replace(latitutde, ",","-"),str_replace(longitude, ",","-")))

```

#remove obsolete columns input_data

```{r}

input_data %<>% select(-c(species, genus))

head(input_data)

```
# add rownumbers for occ ID

```{r}
input_data <- input_data %>% mutate(row_id = row_number())
```

```{r}

input_data %<>%  mutate(dwc_sex = case_when(
                  scientificName == 'Dermacentor reticulatus' ~ tolower(sex),
                  scientificName == 'Ixodes ricinus' ~ tolower(sex),
                  scientificName == 'Ixodes ' ~ tolower(sex)
                  
))



```

```{r}


input_data %<>%  mutate(par_extractiontype = case_when(
                  scientificName == 'borelia' ~ tolower(extractiontype),
                  scientificName == 'tbe' ~ tolower(extractiontype),
                  scientificName == 'babesiacanis_caballi' ~ tolower(extractiontype),
                  scientificName == 'theileriaequi' ~ tolower(extractiontype)
                  
))



```
```




# Occurrence core 


Create dataframes for occurrence (1 ticks, 2)

```{r}
occurrence1 <- input_data


```


# Term mapping occurrence1 -

Map the data to [Darwin Core occurrence](https://rs.gbif.org/core/dwc_occurrence_2020-07-15.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).
# Record-level
# occurrenceID

### occurrenceID

str_c("ITG:FAVV:",Collection.sites_Site.Code))

```{r}
occurrence1 %<>% mutate(dwc_occurrenceID = str_c("ITG:FAVV:IX:OCC:",id, row_id))
```


## eventID



### type

```{r}
occurrence1 %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence1 %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
occurrence1 %<>% mutate(dwc_license = "https://creativecommons.org/licenses/by/4.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence1 %<>% mutate(dwc_rightsHolder = "Institute of Tropical Medicine Antwerp") # e.g. "INBO"
```

### accessRights

```{r}
occurrence1 %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```
### institutionID


### datasetID

```{r}
occurrence1 %<>% mutate(dwc_datasetID = "complete")
```

### institutionCode

```{r}
occurrence1 %<>% mutate(dwc_institutionCode = "ITG") # e.g. "INBO"
```
### collectionCode


```{r}
occurrence1 %<>% mutate(dwc_collectionCode = "ITG:ixodida") # e.g. "INBO"
```

### datasetName

```{r}
occurrence1 %<>% mutate(dwc_datasetName = "FAVV - Exotic vector and pathogen surveillance programme in Belgium - Ixodida_host") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```
### basisOfRecord

```{r}
occurrence1 %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```


```{r}
head(occurrence1, n = 10)
```


# Event

### eventDate

```{r}
occurrence1 %<>% mutate(dwc_eventDate = collectiondate)
   
#paste((Start.Date),"/",(End.Date)                        

```
                           
### samplingProtocol

```{r}
occurrence1 %<>% mutate(dwc_samplingProtocol = 'dragging or by hand collection from vegetation or hosts')
```


### habitat
no habitat available here

```{r eval=FALSE, include=FALSE}
event %<>% mutate(dwc_habitat = pbs_type)
```


# Location

### locationID

```{r}
occurrence1 %<>% mutate(dwc_locationID = str_c("ITG:FAVV:",sitecode)) # e.g. "Belgium = BE"
```

### continent

```{r}
occurrence1 %<>% mutate(dwc_continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
occurrence1 %<>% mutate(dwc_countryCode = "BE") # e.g. "Belgium = BE"
```

### municipality


```{r}
occurrence1 %<>%
  mutate(dwc_municipality = municipality)
```

### locality



```{r}
occurrence1 %<>%
  mutate(dwc_locality = site_name)
```

### decimalLatitude & decimalLongitude

```{r}
head(occurrence1, n = 5)
occurrence1 %<>%
   mutate(decimalLongitude = str_replace(longitude, ",",".")) %>%
   mutate(decimalLatitude = str_replace(latitutde, ",","."))
```


```{r}
occurrence1 %<>%   
   mutate(dwc_decimalLatitude = (round(as.numeric(decimalLatitude), digits = 5))) %>%
   mutate(dwc_decimalLongitude = (round(as.numeric(decimalLongitude), digits = 5)))

head(occurrence1)
   
```

this is not needed anymore here

```{r eval=FALSE, include=FALSE}
event %<>%   
   mutate(dwc_decimalLatitude2 = (round(dwc_decimalLatitude, digits = 5))) %>%
   mutate(dwc_decimalLongitude2 = (round(dwc_decimalLongitude, digits = 5)))
```

### geodeticDatum

```{r}
occurrence1 %<>%
  mutate(dwc_geodeticDatum = 'WGS84')
```

### coordinateUncertaintyInMeters


```{r}
occurrence1 %<>%
  mutate(dwc_coordinateUncertaintyInMeters = '30')
```

### georeferenceRemarks






The following terms contain information about the occurrence:


Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

remove all row not relevant for occurrence (no occurrence)

Identify zero occurrences


```{r eval=FALSE, include=FALSE}
occurrence1[is.na(occurrence)] = 0
```


Won't use, also others than cullicidae here
script to fill empty spaces

```{r eval=FALSE, include=FALSE}
occurrence$Family[occurrence$Family == ""] <- NA_character_
occurrence <- occurrence %>% mutate(dwc_family = replace_na(Family, replace = "Culicidae"))


```


```{r}
occurrence1 %<>%
  mutate(dwc_order = 'Ixodida')
```



Remove all columns not relevant for occurrence

## Remove obsolete columns for occurrence

```{r eval=FALSE, include=FALSE}
occurrence %<>% select(-c(Latitutde,Longitude, Municipality, Collection.sites_Site.Code, Site.Name, Site.type, UnionQuery1_Site.Code, Trap.type, Trap.Code, Start.Date, End.Date, Comments.sampling  ))
```



## eventID


### basisOfRecord

```{r}
occurrence1 %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### recordedBy

```{r}
occurrence1 %<>% mutate(dwc_recordedBy = "ITG")
```

### identifiedBy

```{r}
occurrence1 %<>% mutate(dwc_identifiedBy = 'ITG')
```
### collectionCode

```{r}
occurrence1 %<>% mutate(dwc_collectionCode = "ITG:FAVV")
```


### individualCount

```{r}
occurrence1 %<>% mutate(dwc_individualCount = poolnumber) 
```

### organismQuantity

### organismQuentityType

### sex

```{r}
occurrence1 %<>% mutate(dwc_sex = recode(sex,
                   "egg" = "",
                   "larva"  = "",
                   "pupa" = "",
                   "nymph" = "",
                   "Unknown" = "",
                   "Female" = "female",
                   "Male" = "male",
                   "?" = ""
                                ))
```

### lifeStage

```{r}
occurrence1 %<>% mutate(dwc_lifeStage = recode(sex,
                   "Female" = "adult",
                   "Male"  = "adult",
                   "Unknown" = "",
                   "?" = ""
                                     ))
```
### scientificName

```{r}
occurrence1 %<>% mutate(dwc_scientificName = paste(genus, species))
```




### occurrenceStatus

```{r}
occurrence1 %<>% 
    ##select(individualCount) %>%
    mutate(dwc_occurrenceStatus = case_when(dwc_individualCount > 0 ~ "present",
                              dwc_individualCount == 0 ~ "absent"
                                 )
                                )
```

# identification

### identifiedBy

```{r}
occurrence1 %<>% mutate(dwc_identifiedBy = 'ITG')
```



# taxon

### scientificName 



### kingdom

```{r}
occurrence1 %<>% mutate(dwc_kingdom = "Animalia")
```


# taxonRank

```{r}
occurrence1 %<>% mutate(dwc_taxonRank = case_when(
                   dwc_scientificName == 'Ixodes ricinus' ~ "species",
                   dwc_scientificName == 'Ixodes ' ~ "genus",
                   dwc_scientificName == 'Dermacentor reticulatus' ~ "species"
                                                        ))

#,TRUE ~ "species"
```

# nomenclaturalCode

```{r}
occurrence1 %<>% mutate(dwc_nomenclaturalCode = "ICZN") # e.g. "ICZN"
```


## Post-processing Occurrence



```{r}
occurrence1 %<>% select(starts_with("dwc_"))

colnames(occurrence1) <- str_remove(colnames(occurrence1), "dwc_")
  #remove collection columns
          ##  rename(scientificName = verbatimScientificName) 

occurrence1 %<>% distinct()
```





Preview data:

```{r}
occurrence1 %>% head()
```


Save to CSV:

```{r}
write_csv(occurrence1, here::here("datasets","favv-ixodida", "data", "processed", "occurrence.csv"), na = "")


```


# Data quality tests

```{r}
parsed_names <- occurrence1 %>%
  distinct(scientificName)  %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function

speciesList <- parsed_names %>%
   select(scientificname)

write_csv(speciesList, here::here("datasets","favv-ixodida", "data", "processed", "specieslist.csv"), na = "")
```

