---
title: "dwc_mapping"
author: "Dimitri Brosens"
date: "17-3-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
#library(sf)             # coordinate transformation
library(lubridate)      # date transformation
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was corrected in Excel
Muskrat occurrences opened in openRefine
Obsolete columns removed
some columns renamed to DwC term
File exported to csv

```{r}
#input_interim <- read_csv2(file = here:here("datasets", "mica-uvw-occurrences", "data", "raw", "2020_11_18_Vangstgegevens #muskus- en beverratten_NL_2019.csv"))  

input_data <- read.csv(file = here::here("datasets", "favv-ixodida", "data", "interim", "Integrated_output_ticks_ITG.csv"), sep = ',')
```

Preview data:

```{r}
input_data %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% remove_empty("rows")
```



delete obsolete columns

```{r}
input_data %<>% select(-c("Male","Female","Larva","Nymph","ID.1"))
```
fix issue with decimal lat and decimal long

Not needed in this dataset

```{r eval=FALSE, include=FALSE}
integer_part_lat <- map_chr(input_data$poe_latitude, function(x) str_split(x, pattern = "\\.")[[1]][1])

decimal_part_lat <- map(input_data$poe_latitude, function(x) {
 decimal_parts_lat <- NA_character_
 splitted_string_lat <- str_split(x, pattern = "\\.")[[1]]
 if (length(splitted_string_lat) > 1) {
   decimal_parts_lat <- paste0(splitted_string_lat[2:length(splitted_string_lat)])
   paste0(decimal_parts_lat, collapse = "")
   }
 else {
   ""
 }
})

integer_part_long <- map_chr(input_data$poe_longitude, function(x) str_split(x, pattern = "\\.")[[1]][1])

decimal_part_long <- map(input_data$poe_longitude, function(x) {
 decimal_parts_long <- NA_character_
 splitted_string_long <- str_split(x, pattern = "\\.")[[1]]
 if (length(splitted_string_long) > 1) {
   decimal_parts_long <- paste0(splitted_string_long[2:length(splitted_string_long)])
   paste0(decimal_parts_long, collapse = "")
   }
 else {
   ""
 }
})


input_data %<>%
 mutate(decimalLatitude_poe = as.numeric(paste(integer_part_lat, decimal_part_lat, sep = "."))) %>%
 mutate(decimalLongitude_poe = as.numeric(paste(integer_part_long, decimal_part_long, sep = ".")))
```



```{r}
input_data %>% head(n = 5)
```

# Georeference source data
data is already in WGS84

# Create eventID


```{r}
input_data %<>% mutate(dwc_eventID = str_c("IGT:FAVV:T:EV:",str_replace(Collectiondate,",","-") ,str_replace(Extractiondate, ",","-") ,str_replace(Latitutde, ",","-"),str_replace(Longitude, ",","-")))

```

# Occurrence core 


Create dataframes for occurrence (1 ticks, 2)

```{r}
occurrence1 <- input_data
occurrence2 <- input_data
occurrence3 <- input_data
occurrence4 <- input_data
occurrence5 <- input_data

```


# Term mapping

Map the data to [Darwin Core Event](https://rs.gbif.org/core/dwc_event_2016_06_21.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).
# Record-level
# Event 
## Remove obsolete columns for event

```{r}
event %<>% select(-c(ID,UnionQuery1_Site.Code,Family, Genus, Species, Sex, Number, Box, Storage.of.Box, Status, Identifier, Date.ID, Comments.identification, Trap.Code, Comments.sampling,  Larval.habitat.code, Larval.habitat.type))
```
#group By for eventCore

```{r}
event %<>% distinct()

```



## eventID

add row number

```{r}
event <- event %>% mutate(rowtID = row_number())
```




### type

```{r}
event %<>% mutate(dwc_type = "Event")
```

### language

```{r}
event %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
event %<>% mutate(dwc_license = "https://creativecommons.org/licenses/by/4.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
event %<>% mutate(dwc_rightsHolder = "Institute of Tropical Medicine Antwerp") # e.g. "INBO"
```

### accessRights

```{r}
event %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```
### institutionID


### datasetID

```{r}
event %<>% mutate(dwc_datasetID = "complete")
```

### institutionCode

```{r}
event %<>% mutate(dwc_institutionCode = "ITG") # e.g. "INBO"
```
### collectionCode


```{r}
event %<>% mutate(dwc_collectionCode = "ITG:Mosquitoes") # e.g. "INBO"
```

### datasetName

```{r}
event %<>% mutate(dwc_datasetName = "FAVV - Exotic vector and pathogen surveillance programme in Belgium - Mosquitoes") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```
### basisOfRecord

```{r}
event %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```


```{r}
head(event, n = 10)
```


# Event

### eventDate

```{r}
 event %<>% mutate(dwc_eventDate = paste((Start.Date),"/",(End.Date)))
```
### samplingProtocol

```{r}
event %<>% mutate(dwc_samplingProtocol = Trap.type)
```


### habitat
no habitat available here

```{r eval=FALSE, include=FALSE}
event %<>% mutate(dwc_habitat = pbs_type)
```


# Location

### locationID

```{r}
event %<>% mutate(dwc_locationID = str_c("ITG:FAVV:",Collection.sites_Site.Code)) # e.g. "Belgium = BE"
```

### continent

```{r}
event %<>% mutate(dwc_continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
event %<>% mutate(dwc_countryCode = "BE") # e.g. "Belgium = BE"
```

### municipality


```{r}
event %<>%
  mutate(dwc_municipality = Municipality)
```

### locality



```{r}
event %<>%
  mutate(dwc_locality = Site.Name)
```

### decimalLatitude & decimalLongitude

```{r}
head(event, n = 5)
event %<>%
   mutate(dwc_decimalLongitude = str_replace(Longitude, ",",".")) %>%
   mutate(dwc_decimalLatitude = str_replace(Latitutde, ",","."))
```


```{r}
event %<>%   
   mutate(dwc_decimalLatitude = (round(as.numeric(dwc_decimalLatitude), digits = 5))) %>%
   mutate(dwc_decimalLongitude = (round(as.numeric(dwc_decimalLongitude), digits = 5)))


   
```
this is not needed anymore here
```{r eval=FALSE, include=FALSE}
event %<>%   
   mutate(dwc_decimalLatitude2 = (round(dwc_decimalLatitude, digits = 5))) %>%
   mutate(dwc_decimalLongitude2 = (round(dwc_decimalLongitude, digits = 5)))
```

### geodeticDatum

```{r}
event %<>%
  mutate(dwc_geodeticDatum = 'WGS84')
```

### coordinateUncertaintyInMeters


```{r}
event %<>%
  mutate(dwc_coordinateUncertaintyInMeters = '30')
```

### georeferenceRemarks

## Post-processing Event

```{r}
event %<>% select(starts_with("dwc_"))

colnames(event) <- str_remove(colnames(event), "dwc_")
  #remove collection columns
          ##  rename(scientificName = verbatimScientificName) 

event %<>% distinct()
```

Save to CSV:

```{r}
write_csv(event, here::here("datasets/favv-mosquitoes", "data", "processed", "event.csv"), na = "")



```





The following terms contain information about the occurrence:


Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

remove all row not relevant for occurrence (no occurrence)

Identify zero occurrences


```{r}
occurrence[is.na(occurrence)] = 0
```


Won't use, also others than cullicidae here
script to fill empty spaces

```{r eval=FALSE, include=FALSE}
occurrence$Family[occurrence$Family == ""] <- NA_character_
occurrence <- occurrence %>% mutate(dwc_family = replace_na(Family, replace = "Culicidae"))


```
```{r}
occurrence %<>%
  mutate(dwc_order = 'Diptera')
```



Remove all columns not relevant for occurrence

## Remove obsolete columns for occurrence

```{r}
occurrence %<>% select(-c(Latitutde,Longitude, Municipality, Collection.sites_Site.Code, Site.Name, Site.type, UnionQuery1_Site.Code, Trap.type, Trap.Code, Start.Date, End.Date, Comments.sampling  ))
```
#materialSampleID

```{r}
occurrence %<>% mutate(dwc_materialSampleID = Box)
```


### occurrenceID

str_c("ITG:FAVV:",Collection.sites_Site.Code))

```{r}
occurrence %<>% mutate(dwc_occurrenceID = str_c("ITG:FAVV:OCC:",ID))
```
## eventID


### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = "ITG")
```

### identifiedBy

```{r}
occurrence %<>% mutate(dwc_identifiedBy = Identifier)
```
### collectionCode

```{r}
occurrence %<>% mutate(dwc_collectionCode = "ITG:FAVV")
```


### individualCount

```{r}
occurrence %<>% mutate(dwc_individualCount= Number) 
```

### organismQuantity

### organismQuentityType

### sex

```{r}
occurrence %<>% mutate(dwc_sex = recode(Sex,
                   "egg" = "",
                   "larva"  = "",
                   "pupa" = "",
                   "Unknown" = "",
                   "Female" = "female",
                   "Male" = "male"
                                ))
```

### lifeStage

```{r}
occurrence %<>% mutate(dwc_lifeStage = recode(Sex,
                   "Female" = "adult",
                   "Male"  = "adult",
                   "Unknown" = ""
                                     ))
```
### scientificName

```{r}
occurrence %<>% mutate(dwc_scientificName = paste(Genus, Species))
```


### establishmentMeans

```{r}
occurrence %<>% mutate(dwc_establishmentMeans = recode(dwc_scientificName,
                     "Culicidae" = "" ,
                     "Anopheles claviger" = "native" ,
                     "Anopheles spp." = "native" ,
                     "Culex pipiens/torrentium" = "native" ,
                     "Culiseta annulata" = "native" ,
                     "Culex spp." = "native" ,
                     "Anopheles plumbeus" = "native" ,
                     "Aedes japonicus" = "introduced" ,
                     "Aedes geniculatus" = "native" ,
                     "Culiseta spp." = "native" ,
                     "Anopheles maculipennis s.l." = "native" ,
                     "Anopheles daciae" = "native" ,
                     "Culex hortensis" = "native" ,
                     "Aedes spp." = "" ,
                     "Aedes koreicus" = "introduced" ,
                     "Culiseta longiareolata" = "introduced" ,
                     "Anopheles maculipennis" = "introduced" ,
                     "Aedes rusticus" = "native" ,
                     "Aedes albopictus" = "introduced" ,
                     "Culex modestus" = "vagrant" ,
                     "Culex torrentium" = "native" ,
                     "Culiseta morsitans" = "native" ,
                     "Aedes annulipes/cantans" = "native" ,
                     "Aedes sticticus" = "native" ,
                     "Coquillettidia richiardii" = "native" ,
                     "Aedes vexans" = "native" ,
                     "Aedes communis" = "native" ,
                     "Aedes cinereus/geminus" = "native" ,
                     "Aedes punctor" = "native" ,
                     "Anopheles pharoensis" = "introduced" ,
                     "Aedes caspius/dorsalis" = "native" ,
                     "Culiseta fumipennis" = "native" ,
                     "Culex territans" = "native" ,
                     "Anopheles atroparvus" = "native" ,
                     "Aedes detritus" = "native" 
                                     ))
```
### degreeOfEstablishment

```{r}
occurrence %<>% mutate(dwc_degreeOfEstablishment = recode(dwc_scientificName,
                     "Culicidae" = "" ,
                     "Anopheles claviger" = "native" ,
                     "Anopheles spp." = "native" ,
                     "Culex pipiens/torrentium" = "native" ,
                     "Culiseta annulata" = "native" ,
                     "Culex spp." = "native" ,
                     "Anopheles plumbeus" = "native" ,
                     "Aedes japonicus" = "established" ,
                     "Aedes geniculatus" = "native" ,
                     "Culiseta spp." = "native" ,
                     "Anopheles maculipennis s.l." = "native" ,
                     "Anopheles daciae" = "native" ,
                     "Culex hortensis" = "native" ,
                     "Aedes spp." = "" ,
                     "Aedes koreicus" = "established" ,
                     "Culiseta longiareolata" = "vagrant" ,
                     "Anopheles maculipennis" = "introduced" ,
                     "Aedes rusticus" = "native" ,
                     "Aedes albopictus" = "uncertain" ,
                     "Culex modestus" = "vagrant" ,
                     "Culex torrentium" = "native" ,
                     "Culiseta morsitans" = "native" ,
                     "Aedes annulipes/cantans" = "native" ,
                     "Aedes sticticus" = "native" ,
                     "Coquillettidia richiardii" = "native" ,
                     "Aedes vexans" = "native" ,
                     "Aedes communis" = "native" ,
                     "Aedes cinereus/geminus" = "native" ,
                     "Aedes punctor" = "native" ,
                     "Anopheles pharoensis" = "vagrant" ,
                     "Aedes caspius/dorsalis" = "" ,
                     "Culiseta fumipennis" = "native" ,
                     "Culex territans" = "native" ,
                     "Anopheles atroparvus" = "native" ,
                     "Aedes detritus" = "native" 
                                     ))
```

### pathway


```{r}
occurrence %<>% mutate(dwc_degreeOfEstablishment = recode(dwc_scientificName,
                  "Culicidae" = "" ,
                  "Anopheles claviger" = "native" ,
                  "Anopheles spp." = "native" ,
                  "Culex pipiens/torrentium" = "native" ,
                  "Culiseta annulata" = "native" ,
                  "Culex spp." = "native" ,
                  "Anopheles plumbeus" = "native" ,
                  "Aedes japonicus" = "established" ,
                  "Aedes geniculatus" = "native" ,
                  "Culiseta spp." = "native" ,
                  "Anopheles maculipennis s.l." = "native" ,
                  "Anopheles daciae" = "native" ,
                  "Culex hortensis" = "native" ,
                  "Aedes spp." = "native" ,
                  "Aedes koreicus" = "established" ,
                  "Culiseta longiareolata" = "established" ,
                  "Anopheles maculipennis" = "native" ,
                  "Aedes rusticus" = "native" ,
                  "Aedes albopictus" = "reproducing" ,
                  "Culex modestus" = "native" ,
                  "Culex torrentium" = "native" ,
                  "Culiseta morsitans" = "native" ,
                  "Aedes annulipes/cantans" = "native" ,
                  "Aedes sticticus" = "native" ,
                  "Coquillettidia richiardii" = "native" ,
                  "Aedes vexans" = "native" ,
                  "Aedes communis" = "native" ,
                  "Aedes cinereus/geminus" = "native" ,
                  "Aedes punctor" = "native" ,
                  "Anopheles pharoensis" = "casual" ,
                  "Aedes caspius/dorsalis" = "native" ,
                  "Culiseta fumipennis" = "native" ,
                  "Culex territans" = "native" ,
                  "Anopheles atroparvus" = "native" ,
                  "Aedes detritus" = "native" 
                                     ))
```


### occurrenceStatus

```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(dwc_occurrenceStatus = case_when(dwc_individualCount > 0 ~ "present",
                              dwc_individualCount == 0 ~ "absent"
                                 )
                                )
```

# identification

### identifiedBy

```{r}
occurrence %<>% mutate(dwc_identifiedBy = Identifier)
```

### dateIdentified

```{r}
occurrence %<>% mutate(dwc_dateIdentified = Date.ID)
```

### identificationRemarks

```{r}
occurrence %<>% mutate(dwc_identificationRemarks = Comments.identification)
```

# taxon

### scientificName 



### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```


# taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = case_when(
                   dwc_scientificName == 'Culex spp.' ~ "genus"
                  ,dwc_scientificName == 'Culiseta spp.' ~ "genus"
                  ,dwc_scientificName == 'Culiseta' ~ "genus"
                  ,dwc_scientificName == 'Culex ' ~ "genus"
                  ,dwc_scientificName == 'Anopheles' ~ "genus"
                  ,dwc_scientificName == 'Culicoides spp.' ~ "genus"
                  ,dwc_scientificName == 'Aedes spp.' ~ "genus"
                  ,dwc_scientificName == 'Aedes' ~ "genus"
                  ,dwc_scientificName == '' ~ ''
                  ,dwc_scientificName == 'Anopheles spp.' ~ "genus"
                  ,dwc_scientificName == 'Coquillettidia' ~ "genus"
                  ,TRUE ~ "species"                   ))
```

# nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN") # e.g. "ICZN"
```
# remove unidenfifiable

```{r}
occurrence <- filter(occurrence,Genus != 'Unidentifiable' ) 
```


## Post-processing Occurrence



```{r}
occurrence %<>% select(starts_with("dwc_"))

colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
  #remove collection columns
          ##  rename(scientificName = verbatimScientificName) 

occurrence %<>% distinct()
```


Preview data:

```{r}
occurrence %>% head()
```


Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets","favv-mosquitoes", "data", "processed", "occurrence.csv"), na = "")


```


# Data quality tests

```{r}
parsed_names <- occurrence %>%
  distinct(scientificName)  %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function

speciesList <- parsed_names %>%
   select(scientificname)

write_csv(speciesList, here::here("datasets","favv-mosquitoes", "data", "processed", "specieslist.csv"), na = "")
```

