---
title: "dwc_mapping"
author: "Dimitri Brosens"
date: "22-3-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # date transformation
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was corrected in Excel
Muskrat occurrences opened in openRefine
Obsolete columns removed
some columns renamed to DwC term
File exported to csv

```{r}
#input_interim <- read_csv2(file = here:here("datasets", "mica-uvw-occurrences", "data", "raw", "2020_11_18_Vangstgegevens #muskus- en beverratten_NL_2019.csv"))  

input_data <- read.csv(file = here::here("datasets/memoplus-occurrences", "data", "raw", "MEMOplusDATABASE.csv"), sep = ';')
```

Preview data:

```{r}
input_data %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% remove_empty("rows")

head(input_data, n = 5)
```


delete obsolete columns

```{r}
input_data %<>% select(-c("phone_contact1","phone_contact2","email_contact1","email_contact2","poe_street","poe_postalcode","contactname1","contactname2","collector_2", "polystyrene_status","polystyrene_next_to_trap","polystyrene_bitemarks","presence_pbstype","poe_street","poe_country","sex_stored1","sex_stored2","sex_stored3"))

head(input_data, n = 5)
```
fix issue with decimal lat and decimal long

```{r eval=FALSE, include=FALSE}
integer_part_lat <- map_chr(input_data$poe_latitude, function(x) str_split(x, pattern = "\\.")[[1]][1])

decimal_part_lat <- map(input_data$poe_latitude, function(x) {
 decimal_parts_lat <- NA_character_
 splitted_string_lat <- str_split(x, pattern = "\\.")[[1]]
 if (length(splitted_string_lat) > 1) {
   decimal_parts_lat <- paste0(splitted_string_lat[2:length(splitted_string_lat)])
   paste0(decimal_parts_lat, collapse = "")
   }
 else {
   ""
 }
})

integer_part_long <- map_chr(input_data$poe_longitude, function(x) str_split(x, pattern = "\\.")[[1]][1])

decimal_part_long <- map(input_data$poe_longitude, function(x) {
 decimal_parts_long <- NA_character_
 splitted_string_long <- str_split(x, pattern = "\\.")[[1]]
 if (length(splitted_string_long) > 1) {
   decimal_parts_long <- paste0(splitted_string_long[2:length(splitted_string_long)])
   paste0(decimal_parts_long, collapse = "")
   }
 else {
   ""
 }
})


input_data %<>%
 mutate(decimalLatitude_poe = as.numeric(paste(integer_part_lat, decimal_part_lat, sep = "."))) %>%
 mutate(decimalLongitude_poe = as.numeric(paste(integer_part_long, decimal_part_long, sep = ".")))
```



```{r}
input_data %>% head(n = 5)
```

# Georeference source data
data is already in WGS84

# Occurrence core

## Pre-processing



save input_data for open refine, as we have values in decimalLongitude and decimalLatitude where '.' are misplaced during previuos corrections in ITG



```{r}
write_csv(input_data, here::here("datasets","memoplus-occurrences", "data", "interim", "memoplusRawData.csv"), na = "")

```

```{r}
interim_data <- read.csv(file = here::here("datasets", "memoplus-occurrences", "data", "interim", "memoplusRawDataRefined.csv"), sep = ',')

head(input_data, n=5)
```
Create a dataframe occurrence data only 

```{r}
occurrence <- interim_data
event <- interim_data
```


# Term mapping

Map the data to [Darwin Core Event](https://rs.gbif.org/core/dwc_event_2016_06_21.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).
# Record-level
# Event 
## Remove obsolete columns for event

```{r}

head (event, n = 5)
event %<>% select(-c(trap_status,year, poe_latitude, poe_longitude, species.f, sex_collected, sex_stored1,sex_stored2, sex_stored3 , number, Photo, trap_code, pbs_code))
head (event, n = 5)
```
## eventID

```{r}
event %<>% mutate(dwc_eventID = str_c("IGT:MEMO:EV", Ã...sampling_code))
```


### type

```{r}
event %<>% mutate(dwc_type = "Event")
```

### language

```{r}
event %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
event %<>% mutate(dwc_license = "http://creativecommons.org/licenses/by/4.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
event %<>% mutate(dwc_rightsHolder = "Institute of Tropical Medicine Antwerp") # e.g. "INBO"
```

### accessRights

```{r}
event %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```
### institutionID


### datasetID

```{r}
event %<>% mutate(dwc_datasetID = "complete")
```

### institutionCode

```{r}
event %<>% mutate(dwc_institutionCode = "ITG") # e.g. "INBO"
```


### datasetName

```{r}
event %<>% mutate(dwc_datasetName = "MEMO - Monitoring exotic mosquitoes in Belgium") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```
### basisOfRecord

```{r}
event %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```


```{r}
head(event, n = 10)
```


# Event

### eventDate

```{r}
 event %<>% mutate(dwc_eventDate = paste(dmy(start_date),dmy(end_date), sep = "/"))
```
### samplingProtocol

```{r}
event %<>% mutate(dwc_samplingProtocol = trap_type)
```


### habitat

```{r}
event %<>% mutate(dwc_habitat = pbs_type)
```


# Location

### locationID

```{r}
event %<>% mutate(dwc_locationID = str_c("ITG:MEMO:",poe_code)) # e.g. "Belgium = BE"
```

### continent

```{r}
event %<>% mutate(dwc_continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
event %<>% mutate(dwc_countryCode = "BE") # e.g. "Belgium = BE"
```

### municipality


```{r}
event %<>%
  mutate(municipality = poe_city)
```

### locality

```{r}
event %<>%
  mutate(municipality = poe_name_extra)
```

### decimalLatitude & decimalLongitude

```{r}
head(event, n = 5)
event %<>%
   mutate(dwc_decimalLongitude = round(decimalLongitude_poe, digits = 5)) %>%
   mutate(dwc_decimalLatitude = round(decimalLatitude_poe, digits = 5))
```


```{r}
event %<>%   
   mutate(dwc_decimalLatitude = as.character(format(dwc_decimalLatitude, nsmall = 5))) %>%
   mutate(dwc_decimalLongitude = as.character(format(dwc_decimalLongitude, nsmall = 5)))


   
```

### geodeticDatum

```{r}
event %<>%
  mutate(dwc_geodeticDatum = 'WGS84')
```

### coordinateUncertaintyInMeters


```{r}
event %<>%
  mutate(dwc_coordinateUncertaintyInMeters = '30')
```

### georeferenceRemarks

## Post-processing Event

```{r}
event %<>% select(starts_with("dwc_"))

colnames(event) <- str_remove(colnames(event), "dwc_")
  #remove collection columns
          ##  rename(scientificName = verbatimScientificName) 

event %<>% distinct()
```

Save to CSV:

```{r}
write_csv(event, here::here("datasets/memo-occurrences", "data", "processed", "event.csv"), na = "")



```





The following terms contain information about the occurrence:


Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).


REmove all columns not relevant for occurrence

## Remove obsolete columns for event

```{r}
occurrence %<>% select(-c(latitude,longitude, trap_status,year, poe_latitude, poe_longitude, trap_code, pbs_code, start_date, end_date, trap_type, pbs_type, number_containers_checked, sampling_remarks, poe_type, poe_code, location, poe_type, poe_city, poe_country, poe_municipality, poe_name, poe_name_extra, decimalLatitude_poe, decimalLongitude_poe ))
```

### occurrenceID

**This need to be fixed**

```{r}
occurrence %<>% mutate(dwc_occurrenceID = unique_id_code)
```
## eventID

```{r}
occurrence %<>% mutate(dwc_eventID = str_c("IGT:MEMO:EV", sampling_code))
```

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = collector_1)
```

### identifiedBy

```{r}
occurrence %<>% mutate(dwc_identifiedBy = executor)
```
### collectionCode

```{r}
occurrence %<>% mutate(dwc_collectionCode = "ITG:Memo")
```


### individualCount

```{r}
occurrence %<>% mutate(dwc_individualCount= number) 
```

### organismQuantity

### organismQuentityType

### sex

```{r}
occurrence %<>% mutate(dwc_sex = recode(sex_collected,
                   "egg" = "",
                   "larva"  = "",
                   "pupa" = "",
                                ))
```

### lifeStage

```{r}
occurrence %<>% mutate(dwc_lifeStage = recode(sex_collected,
                   "female" = "adult",
                   "male"  = "adult",
                                     ))
```





### scientificName 

```{r}

occurrence %<>% mutate(dwc_scientificName = species.f)
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```


# taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = case_when(
                   species.f == 'Culex spp.' ~ "genus"
                  ,species.f == 'Culiseta spp.' ~ "genus"
                  ,species.f == 'Aedes spp.' ~ "genus"
                  ,species.f == '' ~ ''
                  ,species.f == 'Anopheles spp.' ~ "genus"
                  ,TRUE ~ "species"                   ))
```

# nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN") # e.g. "ICZN"
```

### occurrenceStatus

```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(dwc_occurrenceStatus = case_when(dwc_individualCount > 0 ~ "present",
                              dwc_individualCount == 0 ~ "absent"
                                 )
                                )
```

## Post-processing Occurrence



```{r}
occurrence %<>% select(starts_with("dwc_"))

colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
  #remove collection columns
          ##  rename(scientificName = verbatimScientificName) 

occurrence %<>% distinct()
```


Preview data:

```{r}
occurrence %>% head()
```


Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets","memo-occurrences", "data", "processed", "occurrence.csv"), na = "")


```
